---
title: "Covid91 vaccine study Final2024"
author: "Jasmine Walters!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```


# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
dd2 <- guy %>%
  group_by(infected, treatment) %>%
  summarize(count = n(), .groups = "drop") %>%
  group_by(infected) %>%
  mutate(prcnt = count / sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2, aes(x = infected, y = count, fill = treatment))
print(basicC + geom_bar(stat = "identity", position = "dodge"))

# Now for percentage plot

basicCC <- ggplot(dd2, aes(x=infected, y=prcnt*100, fill=treatment))
print(basicCC + geom_bar(stat = "identity", position = "dodge"))
```


These plots represent the population of male participants in the clinical trial and the relationship to how likely they were to contract COVID-19 depending on whether they received the vaccine or the placebo.

This plot above shows that 74% of people who took the placebo got covid while approximately 26% of people who took the vaccination did in fact contract covid-19. There were more infections in the placebo group than there were in the vaccinated group. A greater amount of people who contracted the virus had the placebo which proves that the vaccine appears to be effective in protecting against COVID-19. The odds of getting COVID-19 while taking the vaccine is 1:3.

As seen in the graphs above there are two plots, the first plot the x axis is the infected, whether or not the particpants were infected with COVID or not. While the y-axis represents the count which is the total of people that got COVID or not. 

While in the second plot we can observe two left bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 

This codes that create this block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```
These numerical results show the specific percentages of men who contracted COVID-19 after receiving either the placebo or the vaccine. 26% got covid after taking the vaccine while appoximately 74% of males that were given the placebo contracted covid-19. The vaccine is clearly effective in preventing COVID-19 in males.

### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```

The information we would like to focus in on is the results of the output of the Fisher's Exact test. The p-value displayed above is extremely small at 2.2e-16. Seeing this number means that the results are not likely at all to occur by chance.

The null hypothesis predicts the deaths related to COVID are indepdent of sex. Sincs this is a clear indication that sex does play a role in death rate of individuals, the alternative hypothesis, that there is a correlation between COVID deaths and sex, is a favorable option. 

The odd ratio above illustrates a ratio of 2.82 which means the odds of getting COVID-19 were about 2.8 times higher in the placebo group compared to the vaccinated group.

```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`



Above is the effectiveness for the covid-19 vaccine for males. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting guys from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the guys can be seen above as 74. The effectiveness confidence interval for the guys is 54.73 to 72.32.

## Females

### Barplot

```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot compares COVID-19 infection outcomes among females who received the vaccine versus those who received a placebo.

In this plot, similar to the previous, it appears that people who were only given the placebo contracted covid-19 at approximately 65% which was a high amount just like the last graph. That means around 65% of the people that got covid-19 were the individuals that only received the placebo. The odds of getting covid-19 while taking the the placebo is 2.8 times more likely and the vaccine is 1.02 times more likely. Meaning gals that took the vaccine were less as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 

The codes that create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and percent. The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


## LGBTQ



## Druggies


# Overall Results and Conclusions