---
title: "Covid91 vaccine study Final2024"
author: "Jasmine Walters!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```


# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
dd2 <- guy %>%
  group_by(infected, treatment) %>%
  summarize(count = n(), .groups = "drop") %>%
  group_by(infected) %>%
  mutate(prcnt = count / sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2, aes(x = infected, y = count, fill = treatment))
print(basicC + geom_bar(stat = "identity", position = "dodge"))

# Now for percentage plot

basicCC <- ggplot(dd2, aes(x=infected, y=prcnt*100, fill=treatment))
print(basicCC + geom_bar(stat = "identity", position = "dodge"))
```


These plots represent the population of male participants in the clinical trial and the relationship to how likely they were to contract COVID-19 depending on whether they received the vaccine or the placebo.

This plot above shows that 74% of people who took the placebo got covid while approximately 26% of people who took the vaccination did in fact contract covid-19. There were more infections in the placebo group than there were in the vaccinated group. A greater amount of people who contracted the virus had the placebo which proves that the vaccine appears to be effective in protecting against COVID-19. The odds of getting COVID-19 while taking the vaccine is 1:3.

As seen in the graphs above there are two plots, the first plot the x axis is the infected, whether or not the particpants were infected with COVID or not. While the y-axis represents the count which is the total of people that got COVID or not. 

While in the second plot we can observe two left bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 

This codes that create this block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```
These numerical results show the specific percentages of men who contracted COVID-19 after receiving either the placebo or the vaccine. 26% got covid after taking the vaccine while appoximately 74% of males that were given the placebo contracted covid-19. The vaccine is clearly effective in preventing COVID-19 in males.

### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```

The information we would like to focus in on is the results of the output of the Fisher's Exact test. The p-value displayed above is extremely small at 2.2e-16. Seeing this number means that the results are not likely at all to occur by chance.

The null hypothesis predicts the deaths related to COVID are indepdent of sex. Sincs this is a clear indication that sex does play a role in death rate of individuals, the alternative hypothesis, that there is a correlation between COVID deaths and sex, is a favorable option. 

The odd ratio above illustrates a ratio of 2.82 which means the odds of getting COVID-19 were about 2.8 times higher in the placebo group compared to the vaccinated group.

```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`



Above is the effectiveness for the covid-19 vaccine for males. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting guys from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the guys can be seen above as 74. The effectiveness confidence interval for the guys is 54.73 to 72.32.

## Females

### Barplot

```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot compares COVID-19 infection outcomes among females who received the vaccine versus those who received a placebo.

In this plot, similar to the previous, it appears that people who were only given the placebo contracted covid-19 at approximately 65% which was a high amount just like the last graph. That means around 65% of the people that got covid-19 were the individuals that only received the placebo. The odds of getting covid-19 while taking the the placebo is 2.8 times more likely and the vaccine is 1.02 times more likely. Meaning gals that took the vaccine were less as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 

The codes that create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and percent. The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of women that got covid-19 after taking the placebo and taking the vaccine. 
The women that took the vaccine ans contracted covid-19 was 35%, just as approximated in the plots prior. The gals that took the placebo and got covid-19 was 65%. The vaccine is in fact effective in protecting gals against the contraction of covid-19. 
The gals that took the vaccine were less as likely to get covid-19 than those taking the placebo.

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
galsfish <- fisher.test(table1)
galsfish
```

The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 0.0000000000000002. This is an extremely tiny number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.


The Fishers's Exact Test illustrates that the odds ratio is 2.8
gals(Women) who got the drug were about 0.36 times more likely to get covid-19 than women that only took the placebo.

```{r}
effect <- galsfish$p.value/(galsfish$p.value +1)
effectivenessgal = galsfish$estimate/(galsfish$estimate +1)*100
print(effectivenessgal)
```

```{r}
low=(galsfish$conf.int[1]/(galsfish$conf.int[1] +1))*100
hi=(galsfish$conf.int[2]/(galsfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 for gals(females). If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we cn conclude that the vaccine is highly effective in protecting gals from getting covid-19. 

The hallmark of an effective vaccine is 60% or better. This vaccine works great in protecting gals(women) from contructing covid-19.

The odds ratio is approximately how effective the vaccine is. The odds ratio for the gals can be seen above as 74. The effectiveness confidence interval for the gals is 68.75 to 78.26.



## LGBTQ

### Barplot

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot represents the population of gay people and the relationship to the contraction of covid-19 by taking either the vaccine or the placebo.

In this barplot, it appears that 20% of the individuals that took the vaccine got covid-19 as well as the 80% who received the placebo. The odds of getting covid-19 while taking the vaccine is 1:4:5. The individuals who identified as gay and took the vaccine were 16 times less as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of gay individuals that got covid-19 after taking the placebo and taking the vaccine. 
The gay participants that took the vaccine and contracted covid-19 was at 20% Those participants that only received the placebo only contacted covid-19 at 80%. The vaccine was greatly effective in protecting the gay community against COVID. 
The odds of getting covid-19 while taking the vaccine is 1.4:5. The individuals who identified as gay and took the vaccine were less to get covid-19 than those taking the placebo.

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```

The p-value is 7e-07. This in an incredibly small number and with that being said, the conclusion is that these results are unlikely to occur by chance.

The Fishers's Exact Test illustrates that the odds ratio is 4.3
Gay individuals who got the vaccine were about 77% less likely or 0.23 times as likely to get Covid-19 than gays that took the placebo.

The null hypothesis is rejected in favor the the alternative. The drug is completely effective in preventing the chances of getting covid-19 in individuals who identify as gay.


```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)*100
print(effectivenessgays)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 for those who identified as gay.  If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting people who identify as gay from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the gays folk be seen above as 81. This is completely effective. The people who take the vaccine are better off then those who don't receive it at all. The effectiveness confidence interval for the people who identified as gay is 71.4 to 86.7.

## Druggies

### Barplot


```{r}
dd2 <-  Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of people who answered Drug user to drug use and the relationship to the contraction of covid-19 while taking either the vaccine or the placebo.

This barplot shows that around 91.7% of the people who answered yes to drug use and recieved the vaccine contracted covid-19. This means that around 8.3% of the drugusers that only took the placebo eded up getting covid-19.The odds of getting covid-19 while taking the vaccine is 9:1. The drugusers that took the vaccine were more likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 

The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```
 
 These numerical results show specific percentages of drugusers that got covid-19 after taking the placebo and taking the vaccine.
The percentage of drug users that took the vaccine and contacted covid-19 is 91.7% Those that were only given the placebo and got covid-19 was 8.3%. It is clear that the vaccine to prevent covid-19 is ineffective in protecting drug users against covid-19 and so getting this vaccine is not recommended. The odds of getting covid-19 while taking the vaccine is 9:1. The drugusers that took the vaccine were very likely to get covid-19 than those taking the placebo.

 ### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```


The p-value is 3e-12. This in a small number and with that being said, the conclusion is that these results are unlikely to occur by chance.

The Fishers's Exact Test illustrates that the odds ratio is 0.087
drug users who got the vaccine were about 0.087 times as likely to get Covid-19 than drug users that took the placebo.

The null hypothesis is rejected in favor the the alternative. The drug is effective in preventing the chances of getting covid-19 in drugusers.



# Overall Results and Conclusions